Unknown

### 2025-07-03 ms/ms.qmd
revising conclusions

#### MARKER
These results suggest that developmental plasticity, as opposed to acclimation or adaptation to open habitats, may explain the explain the long-standing observation taht amphistomatous leaves are more common in sunny habitats. Our results add to existing explanations by showing high light intensity *per se* does not increase the benefit of amphistomy, but rather than anatomical and biochemical change associated with higher light intensity modulate AA. We cannot yet attribute specific changes wrought by plasticity, but correlational evidence suggests that changes in leaf anatomy may be important. Species responded consistently by increasing stomatal density, especially on the adaxial surface, increasing stomatal density ratio. Sun leaves also had greater gs, Amax. However, these traits were not associated with change in AA (table/figures).

big implications - if trees were amphi, more more carbon sink or reduced transpiration

>>>>

Our results indicate that developmental plasticity, as opposed to acclimation or adaptation to open habitats, may explain the long-standing observation that amphistomatous leaves are more common in sunny habitats. If correct, this implies that high light intensity *per se* does not increase the benefit of amphistomy, but rather than anatomical and/or biochemical changes associated with higher light intensity modulate AA. We cannot yet attribute specific changes wrought by plasticity, but the correlation with laminar LMA suggests that changes in leaf thickness and/or density mediate the relationship between plasticity and AA. While we found little evidence for native light habitat in driving AA, unexplained constitutuve differences among accessions may be byproducts of adaptive evolutionary divergence. Hence, we cannot yet rule out an important role for environmental adaptation in modulating AA and this will require future research to resolve.

#### MARKER
Amphistomy likely plays a significant, but under appreciated, role in the carbon and water balance of plant groups where this trait is common, such as herbaceous plants in sunny habitats and woody species in arid biomes (cites). If our estimates of AA from the wild tomato clades
big implications 
- amphistomy plays role in carbon/water balance. do equivalent gs
- if trees were amphi, more more carbon sink or reduced transpiration
- comparative studies should consider plasticity
